---
title: "Discussion 3"
author: "Huaiyu Hu"
date: "September 23, 2019"
output: html_document
---
```{r}
library(ggplot2)
```


Today, we will do an exercise about Monte Carlo sampling method (Exercise 3.3 and 4.2 in A First Course in Bayesian Statistical Methods. Peter D. Hoff.)

#####Exercise
Tumor counts: A cancer laboratory is estimating the rate of tumorigenesis in two strains of mice, A and B. They have tumor count data for 10 mice in strain A and 13 mice in strain B. Type A mice have been well studied and information from other laboratories suggests that type A mice have tumor counts that are approximately Poisson-distributed with a mean of 12. Tumor count rates for type B mice are unknown, but type B mice are related to type A mice. The observed tumor counts for the two populations are $y_A = (12, 9, 12, 14, 13, 13, 15, 8, 15, 6)$ and $y_B = (11, 11, 10, 9, 9, 8, 7, 10, 6, 8, 8, 9, 7)$.

#####(1) Find the posterior distributions, means, variances and 95% quantile based confidence intervals for $\theta_A$ and $\theta_B$. Assuming a Poisson sampling distribution for each group and the following prior distribution:
$\theta_A \sim gamma(120, 10), \theta_B \sim gamma(12, 1), p(\theta_A, \theta_B) = p(\theta_A) \times p(\theta_B)$

Hint: Conjugate prior for poisson distribution.

#####(2). Compute and plot the posterior expectation of $\theta_B$ under the prior distribution $\theta_B \sim gamma(12 \times n_0, n_0)$ for each value of $n_0 \in {1, 2, ..., 50}$. Describe what sort of prior beliefs about $\theta_B$ would be necessary in order for the posterior expectation of $\theta_B$ to be closed to that of $\theta_A$.

Hint: Create a series of $n_0$ from 1 to 50 firstly. For each $n_0$, calculate its posterior expectation. If necessary, you can draw the plot between $n_0$ and the posterior expectation.

#####(3). For the prior distribution given in part (1), obtain $Pr(\theta_B < \theta_A | y_A, y_B)$ via Monte Carlo sampling.


#####(4). For a range of values of $n_0$, obtain $Pr(\theta_B < \theta_A | y_A, y_B)$ for $\theta_A \sim gamma(120, 10)$ and $\theta_B \sim gamma(12 \times n_0, n_0)$. Describe how sensitive the conclusions about the event {$\theta_B < \theta_A$} are to the prior distribution on $\theta_B$.

Hint: We can use sapply to apply the same function to all the numbers in $n_0$, but you need to plug in exact values in gamma distributions here.

```{r, warning = FALSE}
exps <- sapply(1:50, function(n0) { 
  mean(rgamma(5000, NA, NA) < rgamma(5000, NA, NA))
})

qplot(1:50, exps, geom = c('point', 'smooth'))
```

#####(5). Find the $Pr(\tilde{y}_B < \tilde{y}_A)$ given prior in part (1).
