---
title: "Discussion 10 Solution"
author: "Huaiyu"
date: "11/19/2019"
output: html_document
---
```{r, warning = FALSE, message = FALSE}
library(rstan)
library(bayesplot)
library(rstanarm)
```

## Baysian Linear Regression
Today we see the example of wine quality with the help of rstan and rstanarm package. The goal is to model wine quality based on bayesian linear regression.

### 1. Data
```{r}
d <- read.delim("winequality-red.csv", sep = ";")
d <- d[!duplicated(d), ] # remove the duplicates
dim(d)
```

There are 11 predictors as "fixed.acidity", "volatile.acidity", "citric.acid" , "residual.sugar", "chlorides", "free.sulfur.dioxide" , "total.sulfur.dioxide", "density", "pH", "sulphates" and "alcohol" while response is "quality".
```{r}
names(d)
```

Before writing stan file, we need to set up dataset as below.
```{r}
ds <- as.data.frame(scale(d))
y <- ds$quality
X <- as.matrix(ds[, 1:11])
stan_data <- list(X = X, K = ncol(X), N = nrow(ds), y = y)
```

### 2. Bayesian Linear Model

Assume data $y \sim N(\alpha + \beta X, \sigma^2)$ with prior $\alpha \sim N(0, 10^2)$, $\beta \sim N(0, 10^2)$ and $\sigma \sim Cauchy(0, 2.5)$.
Please write stan file and name it as linear_model.stan.

And then fit your stan model with iteration 1000 times and 4 chains. 
```{r, echo = T, results = 'hide'}
stan_model <- "~/Documents/BU_2019_Fall/linear_model.stan"
fit <- stan(file = stan_model, data = stan_data, iter = 1000, chains = 4)
```
```{r, echo = T, results = 'hide'}
print(fit)
```

```{r}
posterior <- extract(fit, permuted = FALSE)
mcmc_areas(
  posterior,
  pars = c(paste0("beta[",1:11,"]")),
  prob = 0.8, # 80% intervals
  prob_outer = 0.99, # 99%
  point_est = "mean"
)
```


### 3. Posterior Predictive Model Checking
Compare density and empirical cdf of y repicates with observations y, then have a look at intervals. Is this a good model?
```{r}
posterior <- extract(fit)
yrep <- posterior$y_rep
p1 <- ppc_dens_overlay(y, yrep[sample(nrow(yrep), 50), ])
p2 <- ppc_ecdf_overlay(y, yrep[sample(nrow(yrep), 50), ])
cowplot::plot_grid(p1, p2)
```

```{r}
i <- sample(length(y), 50); 
p1 <- ppc_ribbon(y[i], yrep[, i]); p2_1i <- ppc_intervals(y[i], yrep[, i])
cowplot::plot_grid(p1, p2_1i, ncol = 1)
```

```{r}
summary(lm(y ~ X))
```
You can use some techniques from linear model class, such as feature selection with AIC/BIC, detect whether correlation exists between independent variables, add polynomial terms and so on.

#### 4. Rstanarm Package
In the rstanarm package, the prior must be a call to $R^2$ with its location argument specified or NULL, which would indicate a standard uniform prior for the $R^2$. The choice of prior here is based on design matrix, and more details are here: https://cran.r-project.org/web/packages/rstanarm/vignettes/lm.html.

If you want to specify any of the other types of priors, you'll need to switch to stan_glm. With the default family (gaussian), it will be equivalent to stan_lm. This runs faster than rstan package, and you don't need to write anything about stan file.
```{r, echo = T, results = 'hide'}
post <-
  stan_lm(y ~ X, data = ds, prior = R2(location = 0.5), iter = 1000, chains = 4)
```
```{r}
post
```


```{r}
plot(post, prob = 0.8)
plot(post, plotfun = "acf", par = c("Xfixed.acidity", "Xvolatile.acidity", "Xcitric.acid"))
```


```{r}
pp_check(post, plotfun = "stat", stat = "mean")
pp_check(post, plotfun = "stat_2d", stat = c("mean", "sd"))
```

```{r}
yrep2 <- posterior_predict(post, newdata = ds)
i <- sample(length(y), 50); 
p1 <- ppc_ribbon(y[i], yrep2[, i]); p2_1i <- ppc_intervals(y[i], yrep2[, i])
cowplot::plot_grid(p1, p2_1i, ncol = 1)
```

In the later lecture, you can also use rstanarm package when analyzing generlized linear model (GLM).

There will be no discussion next week. Happy Thanksgiving Day!