## 11.3         
Given the following data:  
```{r, echo=TRUE}  
Data              <-  data.frame(machine = rep(1:6, each=5), 
                               measure = c(83, 92, 92, 46, 67,
                                           117, 109, 114, 104, 87,
                                           101, 93, 92, 86, 67,
                                           105, 119, 116, 102, 116,
                                           79, 97, 103, 79, 92,
                                           57, 92, 104, 77, 100))
J               <- length(unique(Data$machine))
n               <- length(Data$measure)
```  
*Starting points*  
Select 10 $\theta_{j}$ randomly from the $y_{ij}$ sample. 
```{r, echo=TRUE}
theta_start         <- sapply(1:6,function(x) sample(Data$measure[Data$machine==x],
                                                 10, replace=TRUE))
```  
*1. Conditional posterior of $\tau^{2}$ *
```{r,echo=TRUE}
TauHat2       <- function(theta) {
  mu            <- mean(theta)
  Tau2         <- ( 1/(J-1) ) * sum((theta - mu)^2)
  return(Tau2)
} 

Tau_CondPosterior      <- function(theta) {
  Tau2         <- TauHat2(theta) 
  Tau_Post      <- (J - 1) * (Tau2)/rchisq(1,J-1)
  return(Tau_Post)
  }
```  
*2. Conditional posterior of $\sigma^{2}$ *
```{r,echo=TRUE}  
FsigmaHat2   <- function(theta) {
  sigmaHat2   <-  sapply(1:6, function(x) (Data$measure[Data$machine==x] - theta[x])^2)  
  sigmaHat2   <-  (1/n) * sum(unlist(sigmaHat2))
  return(sigmaHat2)
}

Sigma2_CodPosterior      <-  function(theta) {
  Sigma2_Hat      <- FsigmaHat2(theta) 
  Sigma2_Post      <-  (n) * (Sigma2_Hat)/rchisq(1,n)
  return(Sigma2_Post)
  }
```  
*3. Conditional posterior of $\mu$ *
```{r,echo=TRUE}
Mu_Hat          <- function(theta) {
  mean(theta)
} 

Mu_CondPosterior        <- function(theta,tau2) {
  Mu_Hat    <- Mu_Hat(theta)
  rnorm(1,Mu_Hat,sqrt(tau2/J))
}
```  
*4. Conditional posterior of $\theta$*
```{r,echo=TRUE}
Theta_Hat_j     <- function(j,mu,sigma2,tau2) {
  Yavg_j       <- mean(Data$measure[Data$machine==j])
  N_j           <- length(Data$measure[Data$machine==j])
  ( (1/tau2) * mu + (N_j/sigma2) * Yavg_j  ) / ( (1/tau2) + (N_j/sigma2)  )
} 
VTheta_Hat_j   <- function(j,mu,sigma2,tau2) {
  N_j           <- length(Data$measure[Data$machine==j])
  ( 1 ) / ( (1/tau2) + (N_j/sigma2) )
} 
Theta_CondPosterior         <- function(mu,sigma2,tau2) {
  theta         <- NULL 
  for (j in 1:J) {
    T_Hat       <- Theta_Hat_j(j,mu,sigma2,tau2)   
    V_T_Hat     <- VTheta_Hat_j(j,mu,sigma2,tau2)
    theta[j]    <- rnorm(1,T_Hat,sqrt(V_T_Hat)) 
  }
  return(theta)
}
```
** Hierarchical Model: Sample using Gibbs Sampler **
```{r,echo=TRUE}
sims          <- 200
Gibbs_Sampler      <- function(start_pt_seq_No) {
  param         <- 9
  sample_parameters       <- matrix(NA, ncol = param, nrow = sims )
  colnames(sample_parameters)<-  c("theta1", "theta2", "theta3", 
                         "theta4", "theta5", "theta6", 
                         "mu", "sigma2", "tau2")
  sample_parameters[1,1:6]<- theta_start[start_pt_seq_No,]
  sample_parameters[1,9]  <- Tau_CondPosterior(theta_start[start_pt_seq_No,])
  sample_parameters[1,8]  <- Sigma2_CodPosterior(theta_start[start_pt_seq_No,])
  sample_parameters[1,7]  <- Mu_CondPosterior(theta_start[start_pt_seq_No,],sample_parameters[1,9])
   
  for (s in 2:sims) {
    sample_parameters[s,1:6]<- Theta_CondPosterior(sample_parameters[s-1,7],sample_parameters[s-1,8],sample_parameters[s-1,9])
    sample_parameters[s,9]  <- Tau_CondPosterior(sample_parameters[s,1:6])
    sample_parameters[s,8]  <- Sigma2_CodPosterior(sample_parameters[s,1:6])
    sample_parameters[s,7]  <- Mu_CondPosterior(sample_parameters[s,1:6],sample_parameters[s,9])
  }
return(sample_parameters)
#Warm-up
}
sample_parameters           <- lapply(1:10, function(x) Gibbs_Sampler(x))
sample_parameters.1         <- sample_parameters[[1]][101:200, ] 
sample_parameters.combined <- rbind(sample_parameters[[1]][101:200, ], sample_parameters[[2]][101:200, ], sample_parameters[[3]][101:200, ], sample_parameters[[4]][101:200, ], sample_parameters[[5]][101:200, ], sample_parameters[[6]][101:200, ], sample_parameters[[7]][101:200, ], sample_parameters[[8]][101:200, ], sample_parameters[[9]][101:200, ], sample_parameters[[10]][101:200, ])
#Transform the variance in to sd.
sample_parameters.combined[,8:9]   <- sqrt(sample_parameters.combined[,8:9] )

# Quantiles
t(apply(sample_parameters.combined,2, function(x) quantile(x, c(.025,.25,.5,.75,.975))))

```  
** Pooled and Separated Model **
```{r, echo=TRUE}  
Yavg_Pooled   <-  mean(Data$measure)
Var_Pooled      <-  sum( (Data$measure[Data$machine==6] - 
                            mean(Data$measure))^2 )/( length(Data$measure) - 1)
Theta_Posterior_Pooled_6<- rnorm(sims, Yavg_Pooled, sqrt(Var_Pooled))
Yavg_Pooled_6     <-  mean(Data$measure[Data$machine==6]) 
Var_Pooled_6       <-  sum( (Data$measure[Data$machine==6] - 
                            mean(Data$measure[Data$machine==6]))^2 ) / 
                        ( length(Data$measure[Data$machine==6]) - 1)
Theta_Posterior_Sep_6   <- rnorm(sims, Yavg_Pooled_6, sqrt(Var_Pooled_6))
```  

** Posterior distribution of the mean of the quality measurements of the sixth machine. For the hierarchical, pooled and separate models **
```{r, echo=TRUE}
plot(density(sample_parameters.combined[,"theta6"]), col="red", 
     xlab="Mean Measure", 
     ylab="Density", 
     main="Posterior Dist of measure mean: 6th Machine, Three Models")
lines(density(Theta_Posterior_Pooled_6), col="blue")
lines(density(Theta_Posterior_Sep_6), col="green")
legend("topleft",col = c("red","blue", "green"),
       legend=c("Hierarchical","pooled", "separated"),
       lty = c(1,1,1)) 
```

** Predictive distribution for another quality measurement of the sixth machine. **
```{r, echo=TRUE}
y_pred_Hierar <- rnorm(1000, mean=sample_parameters.combined[,"theta6"], sd = sample_parameters.combined[,"sigma2"])
y_pred_Pooled <- rnorm(1000, mean = Theta_Posterior_Pooled_6, sd = sqrt(Var_Pooled))
y_pred_Sep <- rnorm(1000, mean = Theta_Posterior_Sep_6, sd = sqrt(Var_Pooled_6))

plot(density(y_pred_Pooled), col= "blue", 
     xlab="Another New Measure", 
     ylab="Density", 
     main="Predictive Dist new measure: 6th machine, Three Models")
lines(density(y_pred_Hierar), col="red")
lines(density(y_pred_Sep), col="green")
legend("topleft",col = c("red","blue", "green"),
       legend=c("Hierarchical","pooled", "separated"),
       lty = c(1,1,1)) 
```


** Posterior distribution of the mean of the quality measurements of the seventh machine. For the hierarchical, pooled and separate models **
```{r, echo=TRUE}
plot(density(sample_parameters.1[,"mu"]), col="red", xlab="Mean Measure", 
    ylab="Density", main="Posterior Dist: 7th Measure, Three Models")
lines(density(Theta_Posterior_Pooled_6), col="blue")
lines(density(Theta_Posterior_Sep_6), col="green")
legend("topleft",col = c("red","blue", "green"), 
       legend=c("Hierarchical","pooled", "separated"), 
       lty = c(1,1,1)) 
```
